---
title: "19F in Germany"
author: "Paloma Cárcamo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, sf, spdep, INLA, cowplot)
```

Missing zipcode: 5685/39315

Shapefiles from GADM https://gadm.org/download_country.html

Vaccination coverage estimates from Robert Kock Institut https://robert-koch-institut.github.io/Inanspruchnahme_von_Routineimpfungen_in_Deutschland-Ergebnisse_aus_der_KV-Impfsurveillance/

```{r}
ger_ipd_raw <- read_csv("data/DE_IPD_deidentified.csv")

ger_ipd <- ger_ipd_raw |> 
  mutate(date = as.Date(DateOfIsolation, format = "%d/%m/%Y")) |> 
  filter(!is.na(PLZpatient) & date > as.Date("2009-01-01")) |> 
  mutate(plz = str_pad(PLZpatient, side = "left", pad = "0", width = 3)) |> 
  group_by(plz, Serotype) |> 
  summarise(ipd_st = n()) |> 
  group_by(plz) |> 
  mutate(ipd = sum(ipd_st)) |> 
  ungroup() |> 
  filter(Serotype == "19F") |> 
  mutate(ratio_19F = ipd_st/ipd)

ger_shp <- read_sf("data/gadm41_DEU_2.shp")

ger_shp_plz <- read_sf("data/plz-2stellig.shp") |> 
  st_transform(st_crs(ger_shp)) |> 
  group_by(plz) |> 
  summarise(geometry = st_union(geometry))

ger_cov <- readxl::read_xlsx("data/kvis_vax.xlsx")

ger_key_raw <- read_csv("data/zuordnung_plz_ort.csv")

ger_key <- ger_key_raw |> 
  mutate(ags2 = substr(ags, 1, 5),
         plz2 = substr(plz, 1, 2)) |> 
  distinct(ags2, plz2)
```

```{r}
ger_covsf <- ger_shp |> 
  right_join(ger_cov, by = c("CC_2" = "county_ags")) |> 
  filter(vaccine == "Pneumokokken" & age_group == "24 Monate")
```

```{r, fig.width = 14, fig.height = 10}
ger_covsf |> 
  ggplot() +
  geom_sf(aes(fill = vacc_rate), color = NA) +
  scale_fill_viridis_c() +
  facet_wrap(~year_birth) +
  theme_void() +
  labs(fill = "Vaccine coverage", title = "Pneumococcal vaccine coverage by AGS and year of birth")
```

```{r, fig.height = 10}
ger_shp |> 
  ggplot() +
  geom_sf(data = ger_shp_plz, aes(fill = plz), color = NA, show.legend = FALSE) +
  geom_sf(color = "black", fill = NA) +
  theme_void()
```


### Aggregate vax rates to first two digits of plz

(toma infinito tiempo en correr)

```{r}
# ger_shp_j <- ger_shp |>
#   mutate(ags_area = st_area(geometry))
# 
# st_intersection_faster <- function(x, y) {
#   y_indices <- st_intersects(x, y) |>
#     unlist() |>
#     unique() |>
#     sort()
# 
#   y_subset <- y[y_indices, ]
#   st_intersection(x, y_subset)
# }
# 
# intersection_pre <- st_intersection_faster(ger_shp, ger_shp_plz)
# 
# intersection <- intersection_pre |>
#   mutate(intersection_area = st_area(geometry),
#          district_area = ger_shp_j$ags_area[match(intersection_pre$CC_2, ger_shp$CC_2)],
#          overlap_percentage = (intersection_area / district_area) * 100)
# 
# plz_ags_key <- intersection |>
#   st_drop_geometry() |>
#   dplyr::select(CC_2, plz, overlap_percentage) |>
#   arrange(CC_2, plz)
# 
# # write_rds(intersection_pre, "data/interim/intersection_pre.rds")
# # write_rds(intersection, "data/interim/intersection.rds")
# # write_rds(plz_ags_key, "data/interim/plz_ags_key.rds")

plz_ags_key <- read_rds("data/interim/plz_ags_key.rds")
```

```{r}
ger_cov_agg <- ger_cov |> 
  filter(vaccine == "Pneumokokken" & age_group == "24 Monate") |> 
  left_join(plz_ags_key, by = c("county_ags" = "CC_2")) |> 
  mutate(new_pop = pop_weight*overlap_percentage,
         vaxed = new_pop*(vacc_rate/100)) |> 
  group_by(plz, year_birth) |> 
  summarise(vax_pop = sum(vaxed, na.rm = TRUE),
            all_pop = sum(new_pop, na.rm = TRUE)) |> 
  mutate(vacc_rate = as.numeric(vax_pop/all_pop))

ger_cov_sum <- ger_cov_agg |> 
  group_by(plz) |> 
  summarise(vacc_rate = mean(vacc_rate, na.rm = TRUE))
```

### First two digits of plz

```{r}
ger_ipd_plz2 <- ger_ipd_raw |> 
  mutate(date = as.Date(DateOfIsolation, format = "%d/%m/%Y")) |> 
  filter(!is.na(PLZpatient) & date > as.Date("2009-01-01")) |> 
  mutate(plz = str_pad(PLZpatient, side = "left", pad = "0", width = 3),
         plz_2 = substr(plz, 1, 2)) |> 
  group_by(plz_2, Serotype) |> 
  summarise(ipd_st = n()) |> 
  group_by(plz_2) |> 
  mutate(ipd = sum(ipd_st)) |> 
  ungroup() |> 
  filter(Serotype == "19F") |> 
  mutate(ratio_19F = ipd_st/ipd)

ger_ipd_plz2_v2 <- ger_ipd_raw |> 
  mutate(date = as.Date(DateOfIsolation, format = "%d/%m/%Y"),
         year = year(date)) |> 
  filter(!is.na(PLZpatient) & date > as.Date("2009-01-01")) |> 
  mutate(plz = str_pad(PLZpatient, side = "left", pad = "0", width = 3),
         plz_2 = substr(plz, 1, 2)) |> 
  group_by(plz_2, Serotype, year) |> 
  summarise(ipd_st = n()) |> 
  group_by(plz_2, year) |> 
  mutate(ipd = sum(ipd_st)) |> 
  ungroup() |> 
  filter(Serotype == "19F") |> 
  mutate(ratio_19F = ipd_st/ipd) |> 
  group_by(plz_2) |> 
  summarise(ratio_19F = mean(ratio_19F, na.rm = TRUE))

ger_ipd_plz2_sf <- ger_shp_plz |> 
  left_join(ger_ipd_plz2, by = c("plz" = "plz_2")) |> 
  mutate(plz = factor(plz)) |> 
  ungroup()

ger_ipd_plz2_sf_v2 <- ger_shp_plz |> 
  left_join(ger_ipd_plz2_v2, by = c("plz" = "plz_2")) |> 
  mutate(plz = factor(plz)) |> 
  ungroup()
```

```{r}
ger_ipd_plz2_sf |> 
  ggplot() +
  geom_sf(aes(fill = ratio_19F), color = NA) +
  scale_fill_viridis_c() +
  theme_void()
   
ger_ipd_plz2_sf_v2 |> 
  ggplot() +
  geom_sf(aes(fill = ratio_19F), color = NA) +
  scale_fill_viridis_c() +
  theme_void()

ger_shp_plz |> 
  right_join(ger_cov_sum, by = "plz") |> 
  ggplot() +
  geom_sf(aes(fill = vacc_rate), color = NA) +
  scale_fill_viridis_c() +
  theme_void()
```

```{r}
ger_shp_plz |> 
  right_join(ger_cov_agg, by = "plz") |> 
  ggplot() +
  geom_sf(aes(fill = vacc_rate), color = NA) +
  scale_fill_viridis_c() +
  facet_wrap(~year_birth) +
  theme_void()
```
### Model setup

```{r}
nb3 <- poly2nb(ger_shp_plz, queen = TRUE)
nb2INLA("map_plz2.adj", nb3)
g_plz2 <- inla.read.graph(filename = "map_plz2.adj")

plz2_db <- ger_shp_plz |> 
  st_drop_geometry() |> 
  ungroup() |> 
  mutate(idarea = row_number()) |> 
  right_join(ger_cov_sum, by = "plz") |> 
  right_join(ger_ipd_plz2, by = c("plz" = "plz_2")) |> 
  mutate(vax_cat = cut(vacc_rate*100, breaks = 20))

data_inla_plz2 <- list(
  cases_19F = plz2_db$ipd_st,
  total_cases = plz2_db$ipd,
  N = plz2_db$ipd,
  idarea = plz2_db$idarea,
  vacc_rate = plz2_db$vacc_rate*100,
  vax_cat = plz2_db$vax_cat
)
```

#### No vaccine coverage

```{r}
f1 <- cases_19F ~ 1 + f(idarea, model = "bym2", graph = g_plz2)

result1 <- inla(
  f1, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result1)

bym_st_plz21 <- ger_ipd_plz2_sf |> 
  cbind(result1$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz21 <- ger_ipd_plz2_sf |> 
  cbind(result1$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

plz2_structured1 <- bym_st_plz21 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured1 <- bym_unst_plz21 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured1, plz2_unstructured1, ncol = 2)
```

#### Linear vaccine coverage

```{r}
f2 <- cases_19F ~ 1 + vacc_rate + f(idarea, model = "bym2", graph = g_plz2)

result2 <- inla(
  f2, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result2)

bym_st_plz22 <- ger_ipd_plz2_sf |> 
  cbind(result2$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz22 <- ger_ipd_plz2_sf |> 
  cbind(result2$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

plz2_structured2 <- bym_st_plz22 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured2 <- bym_unst_plz22 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured2, plz2_unstructured2, ncol = 2)
```

#### IID vaccine coverage (no extra constraint)

```{r}
f3 <- cases_19F ~ 1 + f(vax_cat, model = "iid") + f(idarea, model = "bym2", graph = g_plz2)

result3 <- inla(
  f3, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result3)

bym_st_plz23 <- ger_ipd_plz2_sf |> 
  cbind(result3$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz23 <- ger_ipd_plz2_sf |> 
  cbind(result3$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat3 <- result3$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat3 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured3 <- bym_st_plz23 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured3 <- bym_unst_plz23 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured3, plz2_unstructured3, ncol = 2)
```

#### RW1 vaccine coverage (no extra constraint)

```{r}
f4 <- cases_19F ~ 1 + f(vax_cat, model = "rw1") + f(idarea, model = "bym2", graph = g_plz2)

result4 <- inla(
  f4, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result4)

bym_st_plz24 <- ger_ipd_plz2_sf |> 
  cbind(result4$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz24 <- ger_ipd_plz2_sf |> 
  cbind(result4$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat4 <- result4$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat4 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured4 <- bym_st_plz24 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured4 <- bym_unst_plz24 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured4, plz2_unstructured4, ncol = 2)
```

#### RW2 vaccine coverage (no extra constraint)

```{r}
f5 <- cases_19F ~ 1 + f(vax_cat, model = "rw2") + f(idarea, model = "bym2", graph = g_plz2)

result5 <- inla(
  f5, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result5)

bym_st_plz25 <- ger_ipd_plz2_sf |> 
  cbind(result5$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz25 <- ger_ipd_plz2_sf |> 
  cbind(result5$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat5 <- result5$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat5 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured5 <- bym_st_plz25 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured5 <- bym_unst_plz25 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured5, plz2_unstructured5, ncol = 2)
```
#### IID vaccine coverage (forcing reference: lowest vax)

```{r}
extraconstr <- list(
  A = matrix(0, nrow = 1, ncol = nlevels(plz2_db$vax_cat)),
  e = 0
)

extraconstr$A[1, 1] <- 1

f6 <- cases_19F ~ 1 + f(vax_cat, model = "iid", constr = FALSE, extraconstr = extraconstr) + f(idarea, model = "bym2", graph = g_plz2)

result6 <- inla(
  f6, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result6)

bym_st_plz26 <- ger_ipd_plz2_sf |> 
  cbind(result6$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz26 <- ger_ipd_plz2_sf |> 
  cbind(result6$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat6 <- result6$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat6 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured6 <- bym_st_plz26 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured6 <- bym_unst_plz26 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured6, plz2_unstructured6, ncol = 2)
```

#### RW1 vaccine coverage (forcing reference: lowest vax)

```{r}
f7 <- cases_19F ~ 1 + f(vax_cat, model = "rw1", constr = FALSE, extraconstr = extraconstr) + f(idarea, model = "bym2", graph = g_plz2)

result7 <- inla(
  f7, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result7)

bym_st_plz27 <- ger_ipd_plz2_sf |> 
  cbind(result7$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz27 <- ger_ipd_plz2_sf |> 
  cbind(result7$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat7 <- result7$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat7 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured7 <- bym_st_plz27 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured7 <- bym_unst_plz27 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured7, plz2_unstructured7, ncol = 2)
```

#### RW2 vaccine coverage (forcing reference: lowest vax)

```{r, fig.width = 12}
f8 <- cases_19F ~ 1 + f(vax_cat, model = "rw2", constr = FALSE, extraconstr = extraconstr) + f(idarea, model = "bym2", graph = g_plz2)

result8 <- inla(
  f8, 
  family = "binomial",
  data = data_inla_plz2, 
  Ntrials = data_inla_plz2$N, 
  control.predictor = list(compute = TRUE), 
  control.compute = list(dic = TRUE, waic = TRUE),
  verbose = TRUE
)

summary(result8)

bym_st_plz28 <- ger_ipd_plz2_sf |> 
  cbind(result8$summary.random$idarea[1:95, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

bym_unst_plz28 <- ger_ipd_plz2_sf |> 
  cbind(result8$summary.random$idarea[96:190, ]) |> 
  mutate(sig = case_when(X0.025quant > 0 & X0.975quant > 0 ~ 1,
                         X0.025quant > 0 & X0.975quant < 0 ~ 1,
                         .default = 0))

vaxcat8 <- result8$summary.random$vax_cat |> 
  mutate(ID = factor(ID, levels = levels(plz2_db$vax_cat)))

vaxcat8 |> 
  ggplot(aes(x = ID, color = ID)) +
  geom_errorbar(aes(ymin = `0.025quant`, ymax = `0.975quant`), show.legend = FALSE) +
  geom_point(aes(y = mean), show.legend = FALSE) +
  theme_bw()

plz2_structured8 <- bym_st_plz28 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Structured effects") +
  theme_void()

plz2_unstructured8 <- bym_unst_plz28 |> 
  ggplot() +
  geom_sf(aes(fill = mean), color = NA) +
  scale_fill_viridis_c() +
  labs(title = "Unstructured effects") +
  theme_void()

plot_grid(plz2_structured8, plz2_unstructured8, ncol = 2)
```
