---
title: "Long term trends in IPD in the US"
output: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(viridis)
library(plotly)
library(scales)
library(pheatmap)
#install.packages("waffle", repos = "https://cinc.rud.is") 
library(reshape2)
library(HDInterval)
library(rjags)
library(stringr)
library(tidyr)

```


US ABCs serotype specific IPD data by age

Download the data
```{r, eval=F}
 s1 <- read.csv('https://data.cdc.gov/api/views/qvzb-qs6p/rows.csv?accessType=DOWNLOAD')
 saveRDS(s1,'./Data/ABCs_st_1998_2021.rds') 
```

Create time series by serotype and year

```{r, fig.width=5, fig.height=10}
s1 <- readRDS('./Data/ABCs_st_1998_2021.rds') %>%
  rename(agec = "Age.Group..years.",
         year=Year,
         st=IPD.Serotype,
         N_IPD = Frequency.Count) %>%
  mutate( st= if_else(st=='16','16F', st)) %>%
  group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>%
  ungroup()
```
Heatmap to vizualize trends over entire time period

```{r}

p1 <- s1 %>%
    group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>%
  ggplot(aes(x=year, y=log(N_IPD+0.5), group=st, color=st)) +
  geom_line()+
  theme_minimal()
ggplotly(p1)

#pre PCV7 mean
s2_pre<- s1 %>%
  filter(year %in% c(1998, 1999)) %>%
  group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>%
  ungroup() %>%
    group_by(st) %>%
  summarize(N_IPD_pre=mean(N_IPD)) %>%
  ungroup() %>%
    tidyr::complete(st,  fill=list(N_IPD_pre=0))  #fills 0


s2<- s1 %>%
    group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>% #sum across age group
  ungroup() %>%
    tidyr::complete(st, year, fill=list(N_IPD=0))%>%  #fills 0
  left_join(s2_pre, by='st') %>%
  mutate(N_IPD_pre = if_else(is.na(N_IPD_pre),0, N_IPD_pre) ,
    logRR = log((N_IPD+1)/(N_IPD_pre+1) )) 

max_RR <- s2 %>%
  group_by(st) %>%
  summarize(max_RR = max(logRR)) %>%
  arrange((max_RR))

s2$st <- factor(s2$st, levels = max_RR$st)

df_wide <- s2 %>%
  dplyr::select(year, st, logRR) %>%
  tidyr::pivot_wider(names_from = year, values_from = logRR)

pheatmap(
  df_wide[,-1],
  color = viridis(100),
  cluster_rows = TRUE,
  cluster_cols = F,
  scale = "none",
  border_color = NA,
  labels_row=df_wide$st
)


```

pre-PCV13 heatmap only
```{r, fig.width=5, fig.height=10}

#pre PCV13 mean

s2_pre13<- s1 %>%
  filter(year %in% c(2008, 2009)) %>%
  group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>%
  ungroup() %>%
    group_by(st) %>%
  summarize(N_IPD_pre=mean(N_IPD)) %>%
  ungroup() %>%
    tidyr::complete(st,  fill=list(N_IPD_pre=0))  #fills 0

df_wide <- s2 %>%
  filter(year<=2009) %>%
  dplyr::select(year, st, logRR) %>%
  tidyr::pivot_wider(names_from = year, values_from = logRR)

pheatmap(
  df_wide[,-1],
  color = viridis(100),
  cluster_rows = TRUE,
  cluster_cols = F,
  scale = "none",
  border_color = NA,
  labels_row=df_wide$st
)
```

##post PV13 only

```{r, fig.width=5, fig.height=10}
s2_13<- s1 %>%
    group_by(st, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>% #sum across age group
  ungroup() %>%
    tidyr::complete(st, year, fill=list(N_IPD=0))%>%  #fills 0
  left_join(s2_pre13, by='st') %>%
  mutate(N_IPD_pre = if_else(is.na(N_IPD_pre),0, N_IPD_pre) ,
    logRR13 = log((N_IPD+1)/(N_IPD_pre+1) )) 


df_wide13 <- s2_13 %>%
  dplyr::select(year, st, logRR13) %>%
  filter(year>=2010) %>%
  tidyr::pivot_wider(names_from = year, values_from = logRR13)

pheatmap(
  df_wide13[,-1],
  color = viridis(100),
  cluster_rows = TRUE,
  cluster_cols = F,
  scale = "none",
  border_color = NA,
  labels_row=df_wide13$st
)
```

Trends in PCV7 serotypes over time
```{r}
s1 %>%
  filter(st %in% c('4','6B','9V','14','18C','19F','23F')) %>%
  ggplot(aes(x=year, y=log(N_IPD)))+
  geom_line()+
  facet_wrap(~st, scales ='free') +
  theme_classic()+
  geom_vline(xintercept=c(1999.5, 2009.5), lty=2, color='gray')+
  ggtitle('Trends in PCV7 serotypes')

```

Trends in PCV7 serotypes over time, by Age

```{r, fig.width=14, fig.height=12}
 readRDS('./Data/ABCs_st_1998_2021.rds') %>%
  rename(agec = "Age.Group..years.",
         year=Year,
         st=IPD.Serotype,
         N_IPD = Frequency.Count) %>%
  mutate( st= if_else(st=='16','16F', st),
          agec1 = if_else(agec %in% c("Age <2","Age 2-4") ,1,2 )) %>%
  group_by(st,agec, year) %>%
  summarize(N_IPD=sum(N_IPD)) %>%
  ungroup()%>%
   filter(st %in% c('4','6B','9V','14','18C','19A','19F','23F')) %>%
  ggplot(aes(x=year, y=log(N_IPD+1)))+
  geom_line()+
  facet_grid(agec~st, scales ='free') +
  theme_classic()+
  geom_vline(xintercept=c(1999.5, 2009.5), lty=2, color='gray')+
  ggtitle('Trends in PCV7 serotypes')
```
## Simple JAGS model

This model will fit a simple AR(1) model to data for one serotype

## Prepare the data
```{r}
mod.ds <- s1 %>%
  tidyr::complete(st,year,  fill=list(N_IPD=0)) %>%  #fills 0
  arrange(st, year) %>%
  filter(st== '4') %>%
  group_by(st) %>%
  arrange(year) %>%
  mutate(yearN= row_number()) %>%
  ungroup()
```


### Define the model 
```{r}

model_string_single <- "

model{
for(i in 1:N_years){

  N_IPD[i] ~ dnegbin(prob[i],r)
  prob[i]<- r/(r+lambda[i])  ## likelihood 

  log(lambda[i]) <-  beta1[i]

}

  beta1[1] ~ dnorm(mu, (1 - rho1^2) * tau.beta1) # Prior for the first time point
  for(j in 2:N_years){
    beta1[j] ~ dnorm(mu+ rho1 * beta1[j-1], tau.beta1)
  }

  tau.beta1 ~ dgamma(3, 2)  # Tight prior for tau 
  rho1 ~ dunif(-1, 1)       # Uniform prior for rho

  mu ~ dnorm(0, 1e-4)   # Uninformative prior for intercept
  
  r ~ dunif(0,250)

}
"
```

### Set random seeds. 
We are going to run 3 chains in the MCMC, so we need a seed for each. This ensures reproducible results.

```{r}
inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')
```

### Initialize the model
```{r}
model_spec<-textConnection(model_string_single)
model_jags<-jags.model(model_spec,
                       inits=list(inits1,inits2, inits3),
                       data=list('N_IPD'= mod.ds$N_IPD  ,
                                 'N_years'=max(mod.ds$yearN)
                       ),
                       n.adapt=10000,
                       n.chains=3, quiet=T)
```

### Posterior sampling

what parameters do you want to sample?
```{r}
params<-c('rho1', 'mu','beta1','tau.beta1','lambda')

```

```{r}
posterior_samples<-coda.samples(model_jags, 
                                params, 
                                n.iter=10000)
```

Plot the trace plots

```{r, fig.width=7, fig.height=7}
plot(posterior_samples, ask=T)
```


```{r}
posterior_samples.all<-do.call(rbind,posterior_samples) #combine 3 chains into 1 data frame

post_means<-apply(posterior_samples.all, 2, mean)
sample.labs<-names(post_means)
ci<-t(hdi(posterior_samples.all, credMass = 0.95))
#ci<-matrix(sprintf("%.1f",round(ci,1)), ncol=2)
ci<-matrix(ci, ncol=2)

row.names(ci)<-sample.labs
#post_means<-sprintf("%.1f",round(post_means,1))
names(post_means)<-sample.labs

post_all <- cbind.data.frame(post_means,ci) 
names(post_all) <- c('pred','pred_lcl','pred_ucl')
post_all$param <- row.names(post_all)
```


```{r}

pred.ds <- post_all %>%
  filter(grepl('lambda[',param, fixed=T)) %>%
  mutate(yearN = as.numeric(str_extract(param,  "(?<=\\[)\\d+(?=\\])"))) %>%
  ungroup() %>%
  left_join(mod.ds, by='yearN')

pred.ds%>%
  ggplot(aes(x=yearN ,y=pred)) +
  geom_line() +
  theme_minimal() +
  geom_ribbon(aes(x=yearN, ymin=pred_lcl, ymax=pred_ucl), alpha=0.5) +
  geom_point(aes(x=yearN, y=N_IPD))

```


## Model for all PCV7 serotypes--with globa effects

```{r}
mod.ds.pcv7 <- s1 %>%
  tidyr::complete(st,year,  fill=list(N_IPD=0)) %>%  #fills 0
  arrange(st, year) %>%
  filter(st %in% c('4','6B','9V','14','18C','19F','23F')) %>%
  group_by(st) %>%
  arrange(year) %>%
  mutate(yearN= row_number()) %>%
  ungroup()

mat.pcv7 <- mod.ds.pcv7 %>%
  reshape2::dcast(year~st, value.var='N_IPD') %>%
  dplyr::select(-year)
```

```{r}

model_string_pcv7 <- "

model{
for(i in 1:N_years){
  for(j in 1:N_sts){

  N_IPD[i,j] ~ dnegbin(prob[i,j],r[j])
  prob[i,j]<- r[j]/(r[j]+lambda[i,j])  ## likelihood 

  log(lambda[i,j]) <- epsilon1[i,j] #serotype-specific intercept + AR(1) effect centered around global effects

  }
}

# global AR(1) effect
 beta1[1] ~dnorm( alpha1,(1 - rho_beta^2) * tau.beta1 )
    for(i in 2:N_years){
      beta1[i] ~ dnorm(alpha1 + rho_beta*beta1[i-1], tau.beta1 )
    }

#serotype-specifc AR(1)
  for(j in 1:N_sts){
    epsilon1[1,j] ~ dnorm(delta1[j] + beta1[1], (1 - rho_eps^2) * tau.eps) 
  for(i in 2:N_years){
    epsilon1[i,j] ~ dnorm( delta1[j] + beta1[i] + rho_eps * epsilon1[i-1,j], tau.eps)
  }
  }
  
 ## priors ##
    alpha1 ~ dnorm(0,1e-4)
    tau_global ~dgamma(0.01,0.01)

    rho_beta ~ dunif(-1, 1)       # Uniform prior for rho_beta--global AR(1)
    rho_eps ~ dunif(-1, 1)  #prior for rho_eps same for all STs
    
    tau.beta1 ~ dgamma(3, 2)  # Tight prior for tau, 
    tau.eps ~ dgamma(3, 2)  # Tight prior for tau, shared for all serotypes

    for(j in 1:N_sts){
       delta1[j] ~ dnorm(0, tau_global)  # serotype means centered around 0
       r[j] ~ dunif(0,250) #serotype dispersion parameter
      }
    
}
"
```

```{r}
inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')
```

### Initialize the model
```{r}
model_spec<-textConnection(model_string_pcv7)
model_jags_pcv7<-jags.model(model_spec,
                       inits=list(inits1,inits2, inits3),
                       data=list('N_IPD'= mat.pcv7  ,
                                 'N_years'=max(mod.ds.pcv7$yearN),
                                 'N_sts'=ncol(mat.pcv7)
                       ),
                       n.adapt=10000,
                       n.chains=3, quiet=T)

params<-c('alpha1', 'epsilon1','delta1','beta1','tau.beta1','lambda')

posterior_samples_pcv7 <-coda.samples(model_jags_pcv7, 
                                params, 
                                n.iter=10000)

```

```{r}
posterior_samples.all<-do.call(rbind,posterior_samples_pcv7) #combine 3 chains into 1 data frame

post_means<-apply(posterior_samples.all, 2, mean)
sample.labs<-names(post_means)
ci<-t(hdi(posterior_samples.all, credMass = 0.95))
#ci<-matrix(sprintf("%.1f",round(ci,1)), ncol=2)
ci<-matrix(ci, ncol=2)

row.names(ci)<-sample.labs
#post_means<-sprintf("%.1f",round(post_means,1))
names(post_means)<-sample.labs

post_all <- cbind.data.frame(post_means,ci) 
names(post_all) <- c('pred','pred_lcl','pred_ucl')
post_all$param <- row.names(post_all)
```

```{r}
pcv_sts <- cbind.data.frame('st'= names(mat.pcv7)) %>%
  mutate(stN=row_number())


preds <- post_all %>%
  filter(grepl('lambda', param)) %>%
  mutate(extracted = gsub(".*\\[(.*)\\].*", "\\1", param)
  ) %>%
  separate(extracted, into = c("yearN", "stN"), sep = ",") %>%
  mutate(stN=as.numeric(stN),
         yearN=as.numeric(yearN)) %>%
  left_join(pcv_sts, by='stN') %>%
  left_join(mod.ds.pcv7, by=c('st','yearN')) %>%
  arrange(st, yearN)
```


```{r, fig.width=10, fig.height=8}
preds%>%
  ggplot(aes(x=yearN ,y=pred)) +
  geom_line() +
  theme_minimal() +
  geom_ribbon(aes(x=yearN, ymin=pred_lcl, ymax=pred_ucl), alpha=0.5) +
  geom_point(aes(x=yearN, y=N_IPD)) +
  facet_wrap(~st)
```
